---
title: "programming_challenge_stats"
author: "dani-ol-vel"
date: "2020-09-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
## Programming Challenge Stats
## C. Jacchus

We analyzed different variables to see if we found correlation

```{r read data, echo=FALSE}
# read data
library(utils)
library(tidyverse)
library(dplyr)

#read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", 
                 na.strings = "", fileEncoding = "UTF-8-BOM")

```

## COVID-19 cases and length of a country's name


```{r}
#examine the relationship between name length of countries and covid cases
#mutate to add a variable of name length
data1<- data %>% 
  mutate(name_len= str_length(countriesAndTerritories))

#fit a linear regression model of the variables
model1<-lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ name_len, data1)
summary(model1) 
```

The results suggest significantly negative association between name length and covid cases (p= .023). However, the model is poorly fitted(Adjusted R-squared:  0.0001028). The relationship between name length and covid cases cannot be causal. Based on the Bradford Hill's criteria for causation, a plausible mechanism between cause and effect is important for causal implications. Apparently there lacks a plausibility in the examined relationship.

## Correlation between a weekday and number of reported covid deaths in Poland

```{r}
#convert dates to weekdays in numeric format
data2 <- data
data2$weekday_no <- as.numeric(strftime(as.Date(data2$dateRep, "%d/%m/%Y"), "%u"))

#Subset for Poland
data2_Poland <- data2 %>%
  filter(countriesAndTerritories == "Poland")

#fit a linear regression model
model2 <- lm(deaths ~ weekday_no, data = data2_Poland)
summary(model2)
```

There is a positive significant correlation (p = 0.0408) between the day of a week and number of deaths caused by COVID-19 in Poland. This correlation doesn't seem to occur for many of the other European countries, with focus on Slavic countries, countries with a similar population, as well as countries with similar ratings of anti-COVID response - it seem to be accidental. In other words, in order to obtain this correlation, the country has been cherry-picked from the pool of other countries, that don't show any similar correlation, and the relationship lacks plausible causality.
